Quantifying the importance of an inference model in Bayesian phylogenetics

### 2020-07-26 pirouette_article.tex
Changes @rsetienne

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_30/errors.png}
  \caption{
    \new{The impact of the tree prior for the example tree in Figure 2. The alignment for this true tree was generated using a JC69 substitution model and strict clock model. For inferring the tree from this alignment in the 'generative' scenario the same substitution and clock models were used, and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown). In the 'best' scenario the best-fitting candidate models (JC69 substitution model, relaxed log-normal clock model and BD tree prior) were used. For the twin tree, the same models were chosen, except that BD was used to generate and infer the twin tree in the 'generative' scenario.
    The twin distributions show the baseline inference error.}
    Vertical dashed lines show the median error value per distribution.
  }
  \label{fig:example_30}
\end{figure}

In the upper panel of Figure \ref{fig:example_30},
we can see that the error distributions of the \new{ (assumed) generative model (i.e. the known generative substitution and clock models, and the tree model that is used for generating and inferring the twin tree and assumed in inference of the true tree)}
differ strongly between the true and twin tree. 
This difference shows the extent of the mismatch between
the true tree model (which is DD) and the (Yule) tree prior used in inference.
Because these distributions are distinctively different,
the inference error made when using an 
incorrect (that is, Yule) tree prior on a DD tree
is profound.

@@

The candidate model that had highest evidence given the simulated alignment,
was JC, RLN, BD (see Table \ref{tab:options} for the meaning of these 
abbreviations). JC was indeed the site model used when simulating the
alignments. The RLN clock model is an unexpected choice: the RLN clock
model assumes that mutation rates differ between lineages, whereas the
alignment was simulated with equal mutation rates for all lineages.
The BD model is indeed closer to the DD model than Yule, as both BD and DD
model allow for extinction, while Yule does not.

@@

We showed how to use \verb;pirouette; to quantify the \new{impact} of a 
tree prior in Bayesian phylogenetics, assuming the simplest standard 
tree model possible.
In principle any other standard tree model can be assumed, 
but we chose to provide the simplest example.

Figure~\ref{fig:example_30} illustrates the primary result of our pipeline: 
it shows the error distributions for the true tree and the twin tree \new{
when either the generative model (for substitution and clock models these are known, for the tree model it must be assumed for the true tree and it is known for the twin tree) or the best-fitting set candidate model (i.e. combination of tree model, substitution model and clock model) is used in inference.} 
The clear difference between the error distributions 
for the true tree and the twin tree suggests 
that the choice of tree prior does matter.
We note, however, that only one tree from a novel tree model
is not enough to determine the \new{impact} of using an incorrect
tree prior. Instead, a distribution 
of multiple trees, generated by the novel tree model, should be used. In the supplementary material we have provided some examples.

Like most phylogenetic experiments, the setup of \verb;pirouette;
involves many choices. A prime example is the
length of the simulated DNA sequence. One expects that the inference error
decreases for longer DNA sequences. We investigated this
superficially and confirmed this prediction (see the supplementary materials). 
However, we note that for longer DNA sequences, the assumption 
of the same substitution rates across the entire sequence may become less realistic (different genes may experience different substitution rates)
and hence longer sequences may require more parameters. 
Hence, simply getting longer sequences will not always lead to a drastic 
reduction of the influence of the species tree prior.
Fortunately, \verb;pirouette; provides a pipeline that works for all choices.

Interpreting the results of \verb;pirouette; is up to the user; 
\verb;pirouette; does not answer the question 
whether the inference error is too large to trust the inferred tree. The user is encouraged to use different statistics to measure the error. The nLTT statistic is
a promising starting point, as it can compare any two trees and 
results in an error distribution of known range, but one may also explore other statistics.
In principle, \verb;pirouette; allows for this, but in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
\new{
  For models that make different predictions on topology, 
  the twinning process should be modified.
}
\new{
	As noted in the introduction, Duch\^{e}ne and 
	colleagues [\cite{duchene2018phylodynamic}]
	also developed a method to assess the adequacy of a tree model
	on empirical trees. They simulated trees from the posterior distribution of 
	the parameters and then compared this to the originally inferred tree using 
	tree statistics, to determine whether the assumed tree model in inference 
	indeed generates the tree as inferred. This is useful if these trees match, 
	but when they do not, this does not mean that the inferred tree is incorrect; 
	if sufficient data is available the species tree prior may not be important, 
	and hence the inference may be adequate 
	even though the assumed species tree prior is not. 
	In short, the approach is applied to empirical trees and 
	compares the posterior and prior distribution of trees (with the latter 
	generated with the posterior parameters!).
    By contrast,}
    \verb;pirouette;
	\new{aims to expose when assuming standard priors for the 
	species tree are a mis- or underparameterization. 
	Hence, our approach applies to simulated trees and compares 
	the posterior distributions of trees generated with 
	a standard and non-standard model, but inferred with a standard one. 
	The two methods therefore complement one another.
}
However, we note that the \verb;pirouette; pipeline is not restricted 
to exploring the effects of a new species tree model. 
The pipeline can also be used to explore the effects of non-standard 
clock or site models, such as relaxed clock models with a non-standard 
distribution, correlated substitutions on sister lineages, or elevated 
substitutions rates during speciation events. 
It is, however, beyond the scope of this paper to discuss all these options 
in more detail.

In conclusion, \verb;pirouette; can show the errors to be expected
when the model assumed in inference is different from the 
actual generative model.
The user can then judge whether or not this new model 
should be implemented in our Bayesian phylogenetic tool.

>>>>

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_30/errors.png}
  \caption{
    \new{The impact of the tree prior for the example tree in Figure \ref{fig:dd_tree}. The alignment for this true tree was generated using a JC69 substitution model and strict clock model. For inferring the tree from this alignment in the 'generative' scenario the same substitution and clock models were used, and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown). In the 'best' scenario the best-fitting candidate models (JC69 substitution model, relaxed log-normal clock model and Yule tree prior) were used. For the twin tree, the same models were chosen, except that BD was used to generate and infer the twin tree.
    The twin distributions show the baseline inference error.}
    Vertical dashed lines show the median error value per distribution.
  }
  \label{fig:example_30}
\end{figure}

In the upper panel of Figure \ref{fig:example_30},
we can see that the error distributions of the \new{ (assumed) generative model (i.e. the known generative substitution and clock models, and the tree model that is assumed in inference of the true tree, and the tree model that is used for generating and inferring the twin tree)}
differ strongly between the true and twin tree. 
This difference shows the extent of the mismatch between
the true tree model (which is DD) and the (Yule) tree prior used in inference.
Because these distributions are distinctively different,
the inference error made when using an 
incorrect tree prior on a DD tree is profound.

@@

The candidate model that had highest evidence given the simulated alignment,
was JC69, RLN and Yule (see Table \ref{tab:options} for the meaning of these 
abbreviations). JC69 was indeed the site model used when simulating the
alignments. The RLN clock model is a surprising result: it assumes that mutation rates differ between lineages, whereas the
alignment was simulated with equal mutation rates for all lineages \new{(strict clock)}.
\new{The Yule model is not surprising as it is the standard tree prior that is probably closest to the true DD model because it shows no pull-of-the-present (but also no slowdown).}

@@

We showed how to use \verb;pirouette; to quantify the \new{impact of a 
tree prior in Bayesian phylogenetics, assuming - for illustrative purposes - the simplest standard substitution, clock and
tree models, but also the models that would be selected among many different standard tree priors according to the highest marginal likelihood, as this would be a likely strategy for an empiricist. We recommend exploring different candidate models, but note that this is computationally highly demanding.}

Figure~\ref{fig:example_30} illustrates the primary result of our pipeline: 
it shows the error distributions for the true tree and the twin tree \new{
when either the generative model (for substitution and clock models these are known, for the tree model it must be assumed for the true tree and it is known for the twin tree) or the best-fitting set candidate model (i.e. combination of tree model, substitution model and clock model) is used in inference.} 
The clear difference between the error distributions 
for the true tree and the twin tree suggests 
that the choice of tree prior matters.
We note, however, that only one tree from a novel tree model
is not enough to determine the \new{impact} of using an incorrect
tree prior. Instead, a distribution 
of multiple trees, generated by the novel tree model, should be used. In the supplementary material we have provided some examples.

Like most phylogenetic experiments, the setup of \verb;pirouette;
involves many choices. A prime example is the
length of the simulated DNA sequence. One expects that the inference error
decreases for longer DNA sequences. We investigated this
superficially and confirmed this prediction (see the supplementary material). 
However, we note that for longer DNA sequences, the assumption 
of the same substitution rates across the entire sequence may become less realistic (different genes may experience different substitution rates)
and hence longer sequences may require more parameters. 
Hence, simply getting longer sequences will not always lead to a drastic 
reduction of the influence of the species tree prior.
Fortunately, \verb;pirouette; provides a pipeline that works for all choices.

Interpreting the results of \verb;pirouette; is up to the user; 
\verb;pirouette; does not answer the question 
whether the inference error is too large to trust the inferred tree. The user is encouraged to use different statistics to measure the error. The nLTT statistic is
a promising starting point, as it can compare any two trees and 
results in an error distribution of known range, but one may also explore other statistics.
In principle, \verb;pirouette; allows for this, but in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
\new{
  For models that make different predictions on topology, 
  the twinning process should be modified.
}
\new{
	As noted in the introduction, Duch\^{e}ne and 
	colleagues [\cite{duchene2018phylodynamic}]
	also developed a method to assess the adequacy of a tree model
	on empirical trees. They simulated trees from the posterior distribution of 
	the parameters and then compared this to the originally inferred tree using 
	tree statistics, to determine whether the assumed tree model in inference 
	indeed generates the tree as inferred. This is useful if these trees match, 
	but when they do not, this does not mean that the inferred tree is incorrect; 
	if sufficient data is available the species tree prior may not be important,
	and hence inference may be adequate 
	even though the assumed species tree prior is not. 
	In short, the approach is applied to empirical trees and 
	compares the posterior and prior distribution of trees (with the latter 
	generated with the posterior parameters!).
    By contrast,}
    \verb;pirouette;
	\new{aims to expose when assuming standard priors for the 
	species tree are a mis- or underparameterization. 
	Hence, our approach applies to simulated trees and compares 
	the posterior distributions of trees generated with 
	a standard and non-standard model, but inferred with a standard one. 
	The two methods therefore complement one another.
}
Furthermore, we note that the \verb;pirouette; pipeline is not restricted 
to exploring the effects of a new species tree model. 
The pipeline can also be used to explore the effects of non-standard 
clock or site models, such as relaxed clock models with a non-standard 
distribution, correlated substitutions on sister lineages, or elevated 
substitutions rates during speciation events. 
It is, however, beyond the scope of this paper to discuss all these options 
in more detail.

In conclusion, \verb;pirouette; can show the errors to be expected
when the model assumed in inference is different from the actual generative model.
The user can then judge whether or not this new model 
should be implemented in the Bayesian phylogenetic tool.

### 2020-07-26 pirouette_supplement.tex
Changes @rsetienne

\begin{itemize}
  \item{
    subsection \ref{subsec:guidelines}: guidelines for users
  }
  \item{
    subsection \ref{subsec:installation}: installation
  }
  \item{
    subsection \ref{subsec:resources}: resources, such as 
    website, tutorials, packages used, bug reporting and contributing
  }
  \item{
    subsection \ref{subsec:citation}: citation of pirouette
  }
  \item{
    subsection \ref{subsec:twinning}: the
    twinning process
  }
  \item{
    subsection \ref{subsec:candidates}: candidate models for the inference
  }
  \item{
    subsection \ref{subsec:stochasticity}: the effects of stochasticity
  }
  \item{
    subsection \ref{subsec:nltt}: the
    nLTT statistics
  }
  \item{
    subsection \ref{subsec:main_functions}: main functions
  }
  \item{
    subsection \ref{subsec:main_example}: code, extra figures and
    diagnostics regarding the main example.
  }
  \item{
    subsection \ref{subsec:distribution}: the result of using 
    multiple trees, as generated by the
    same stochastic process as the main example
  }
  \item{
    subsection \ref{subsec:n_taxa}: the effect of the number
    of taxa
  }
  \item{
    subsection \ref{subsec:n_nucleotides}: the effect of the DNA
    alignment sequence length
  }
  \item{
    subsection \ref{subsec:simplest_correct_parameterization} shows the
    effect when doing an inference in the simplest use case
  }
  \item{
    subsection \ref{subsec:under_parameterization} shows the
    effect when doing an inference with an under-parameterization
  }
  \item{
    subsection \ref{subsec:different_n_mutations} shows the
    effect when the twin alignment is allowed to have a different
    number of \new{substitution}s
  }
  \item{
    subsection \ref{subsec:mutation_rate} shows the
    effect of different mutation rates
  }
  \item{
    subsection \ref{subsec:acknowledgments}: Acknowledgments
  }
  \item{
    subsection \ref{subsec:data_accessibility}: Data accessibility
  }
  \item{
    subsection \ref{subsec:author_contributions}: Author contributions
  }
\end{itemize}

@@

Similar to lab experiments, \verb;pirouette; allows to perform
a control measurement, by use of a process we call twinning. 
In this context, this control results in an error distribution
that is the baseline error of the pipeline. The difference
between the 'true' and 'twin' error distributions is caused only
by the mismatch between the (known-to-be-true or assumed) 
generative tree model and the 
tree prior used in the actual inference.

@@

\new{
  We choose to preserve the original topology to 
  increase the similarity between the twin to the original tree. 
  This works well in the cases of BD or DD models 
  we consider in our example, because all these models make the same
  assumption about topology (all topologies are equally likely).
  However, this might not be suited for new models that assign different probabilities
  to trees with the same branching times but different topologies.
}
\new{
  The default option for the twin diversification model $p_T$ 
  is the standard BD model.
}
\verb;pirouette;
\new{
   has a built-in function to use a Yule model as well.
  Additionally, a user can specify a function to generate a twin tree
  from any speciation model, such as, for example, a coalescent model.
}

@@

The second step ($T_2$) of the twinning process simulates the twin alignment 
with the same clock model, site model and mutation rate 
used to simulate the original alignment. 
The twin alignment can be simulated in any user-defined way.
\verb;pirouette; supplies the option to have it simulated with
the same mutation rate as the true alignment. By default, however,
not only the same mutation rate is used, but also the total number of 
\new{substitution}s
matches the true alignment. The total number of \new{substitution}s is defined
as the amount of different nucleotides between the (known) root sequence
compared to the sequences at the tips.

@@

If twinning is used, a candidate model that has the highest evidence for
the twin alignment is also used to create another error
distribution.

@@

The nLTT statistic is the absolute difference
between the normalized lineages-through-time plots of two trees.
The nLTT statistic is chosen, as it can operate on any two trees (regardless
of their crown ages and number of taxa) and its results have a clear range
from zero to one. This normalized result makes it possible to compare trees 
from a distribution of trees from any tree model.
\new{
  The nLTT statistic is unsuitable, however, to distinguish
  between a constant-rate BD model and a family of time-dependent 
  models [\cite{louca2020extant}].
}

@@

From the 'true alignment' the generative inference model is run.
Of course, it cannot be the actual (DD) model. Instead, the
default BEAST2 inference model is used, which assumes a JC69 nucleotide
substitution model, a strict clock model and a Yule tree model.
The resulting posterior trees are shows in the
center-left panel of figure \ref{fig:pipeline}.

From this 'generative true' posterior (center-left panel in 
figure \ref{fig:pipeline}), the difference between each of its trees is
compared to the 'true tree' (top-left panel), using the nLTT statistic,
resulting in the error distribution shown at the bottom-left panel
of figure \ref{fig:pipeline}.

Based on the 'true alignment' (center-left panel), the candidate model
with the highest marginal likelihood is determined, from a set of 
15 models. The set of models consists out of all combinations
of all 4 nucleotides substitution models (JC69, HKY, TN93, GTR),
all 2 clock models (strict and relaxed log-normal) and 2 birth-death
models (Yule and Birth-Death), except for the inference model used as the
generative model (JC69, strict, Yule). The inference model that had the
highest evidence (as shown in Table \ref{tab:true_evidence}) was 
the inference model with a JC69 nucleotide substitution model,
a relaxed log-normal clock model and a Yule tree model.
From the 'true alignment' this best candidate inference model is run.
The resulting posterior trees are shown in the second panel
of the third row of posteriors in figure \ref{fig:pipeline}.

From this 'best true' posterior (second panel in third row of
figure \ref{fig:pipeline}), the difference between each of its trees is
compared to the 'true tree' (top-left panel), using the nLTT statistic,
resulting in the second error distribution in the bottom row
of figure \ref{fig:pipeline}.

@@

From this 'twin tree', a 'twin alignment' is simulated, using
the JC69 nucleotide substitution model and a strict clock model.
The resulting alignment is shown at the center-right
of figure \ref{fig:pipeline}.

From the 'twin alignment' the generative inference model is run as well.
Also here, the default BEAST2 inference model is used, 
which assumes a JC69 nucleotide
substitution model, a strict clock model and a Yule tree model.
The resulting posterior trees are shows in the
third panel of the third row of figure \ref{fig:pipeline}.

From this 'generative twin' posterior (third panel in 
third row of figure \ref{fig:pipeline}), 
the difference between each of its trees is
compared to the 'twin tree' (top-right panel), using the nLTT statistic,
resulting in the error distribution shown at the third panel
of the bottom row of figure \ref{fig:pipeline}.

Based on the 'twin alignment' (center-right panel), the candidate model
with the highest marginal likelihood is determined, from the same 
set of 15 candidate models.The inference model that had the
highest evidence (as shown in Table \ref{tab:twin_evidence}) was 
the inference model with a JC69 nucleotide substitution model,
a strict clock model and a BD tree model (note: this indeed matches how
the twin tree and twin tree have been simulated).
From the 'twin alignment' this best candidate inference model is run.
The resulting posterior trees are shown in the fourth panel
of the third row of posteriors in figure \ref{fig:pipeline}.

From this 'best twin' posterior (fourth in third row of
figure \ref{fig:pipeline}), the difference between each of its trees is
compared to the 'twin tree' (top-right panel), using the nLTT statistic,
resulting in the fourth error distribution in the bottom row
of figure \ref{fig:pipeline}.

@@

\new{
  From figures \ref{fig:example_500_nucleotides}, 
  \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} 
  we can observe that the discrepancy between the true and twin error 
  distributions tend to become smaller as the number of nucleotides increase (see also Fig. \ref{fig:error_vs_seqlength}). 
  This occurred for both the generative and best candidate cases.
  This follows the expectation that a prior becomes less important when
  more information becomes available.
}

@@

\new{
  Since the two models are very similar to each other (the BD model can be 
  turned into a Yule model just by setting the extinction parameter to 
  zero [\cite{nee1994reconstructed}]) the median discrepancy is
  almost negligible. 
  However, with respect to the previous 
  case (subsection \ref{subsec:simplest_correct_parameterization}), 
  where a Yule tree prior has been used, 
  the distributions here exhibit a greater difference.
}
\new{As we use only extant trees, it is reasonable that the method
  is slightly weaker in distinguishing between the Yule and BD
  models. It is unknown what the discriminatory power would be
  when comparing trees with extinction events.}

@@

\new{
  Here the median errors are similar in the three settings and similar to 
  ones relative to the full dataset of \ref{subsec:distribution}. We can also 
  notice that in the case of median likelihood, the twin median error appear 
  to be lower than the true mean error. This usually is a sign that the number 
  of replicates (in this case 10) is too low to allow us to draw precise 
  conclusions from this test. We decided to not explore further in this 
  direction using more simulations because computational times turned to be 
  extremely high. 
  The entire run took 120 hours in total.
}

>>>>

\begin{itemize}
  \item{
    subsection \ref{subsec:guidelines}: guidelines for users
  }
  \item{
    subsection \ref{subsec:installation}: installation
  }
  \item{
    subsection \ref{subsec:resources}: resources, such as 
    website, tutorials, packages used, bug reporting and contributing
  }
  \item{
    subsection \ref{subsec:citation}: citation of pirouette
  }
  \item{
    subsection \ref{subsec:twinning}: the
    twinning process
  }
  \item{
    subsection \ref{subsec:candidates}: candidate models for the inference
  }
  \item{
    subsection \ref{subsec:stochasticity}: the effects of stochasticity
  }
  \item{
    subsection \ref{subsec:nltt}: the
    nLTT statistic
  }
  \item{
    subsection \ref{subsec:main_functions}: main functions
  }
  \item{
    subsection \ref{subsec:main_example}: code, extra figures and
    diagnostics regarding the main example.
  }
  \item{
    subsection \ref{subsec:distribution}: the result of using 
    multiple trees, as generated by the
    same stochastic process as the main example
  }
  \item{
    subsection \ref{subsec:n_taxa}: the effect of the number of taxa
  }
  \item{
    subsection \ref{subsec:n_nucleotides}: the effect of the DNA
    alignment sequence length
  }
  \item{
    subsection \ref{subsec:simplest_correct_parameterization} shows the
    effect when performing inference in the simplest use case
  }
  \item{
    subsection \ref{subsec:under_parameterization} shows the
    effect when performing inference with an under-parameterization
  }
  \item{
    subsection \ref{subsec:different_n_mutations} shows the
    effect when the twin alignment is allowed to have a different
    number of \new{substitution}s
  }
  \item{
    subsection \ref{subsec:mutation_rate} shows the
    effect of different mutation rates
  }
  \item{
    subsection \ref{subsec:acknowledgments}: Acknowledgments
  }
  \item{
    subsection \ref{subsec:data_accessibility}: Data accessibility
  }
  \item{
    subsection \ref{subsec:author_contributions}: Author contributions
  }
\end{itemize}

@@

\verb;pirouette; allows to perform
a control measurement, by use of a process we call twinning. This control results in an error distribution
that is the baseline error of the pipeline. The difference
between the 'true' and 'twin' error distributions is caused only by the mismatch between the true tree model and the 
tree prior used in the actual inference.

@@

\new{
  We chose to preserve the original topology to 
  increase the similarity between the twin to the original tree. 
  This works well in the cases of BD or DD models 
  we consider in our example, because all these models make the same
  assumption about topology (all topologies are equally likely).
  However, this might not be suitable for new models that assign different probabilities
  to trees with the same branching times but different topologies.
}
\new{
  The default option for the twin diversification model $p_T$ 
  is the standard BD model.
}
\verb;pirouette;
\new{
   has a built-in function to use a Yule model as well.
  Additionally, a user can specify a function to generate a twin tree
  from any speciation model, such as, for example, a coalescent model.
}

@@

The second step ($T_2$) of the twinning process simulates the twin alignment 
with the same clock model, site model and mutation rate 
used to simulate the alignment on the true. 
The twin alignment can be simulated in any user-defined way.
\verb;pirouette; provides the option simulate it with the same mutation rate as the true alignment. By default, however,
not only the same mutation rate is used, but also the total number of 
\new{substitution}s
matches the true alignment. The total number of \new{substitution}s is defined
as the number of different nucleotides between the (known) root sequence
compared to the sequences at the tips.

@@

If twinning is used, a candidate model that has the highest evidence for
the twin alignment is also used to create the twin error
distribution.

@@

The nLTT statistic is the absolute difference
between the normalized lineages-through-time plots of two trees.
The nLTT statistic is chosen, as it can operate on any two trees (regardless
of their crown ages and number of taxa) and its results have a clear range
from zero to one. This normalized result makes it possible to compare trees 
from a distribution of trees from any tree model.
\new{
  The nLTT statistic is not suitable, however, to distinguish
  between a constant-rate BD model and a family of time-dependent 
  models [\cite{louca2020extant}].
}

@@

From the 'true alignment' the generative inference model is run.
Of course, it cannot be the actual (DD) model. Instead, the
default BEAST2 inference model is used, which assumes a JC69 nucleotide
substitution model, a strict clock model and a Yule tree model.
The resulting posterior trees are shown in the
center-left panel of figure \ref{fig:pipeline}.

From this 'generative true' posterior (center-left panel in figure \ref{fig:pipeline}), the difference between each of its trees is
compared to the 'true tree' (top-left panel), using the nLTT statistic,
resulting in the error distribution shown in the bottom-left panel
of figure \ref{fig:pipeline}.

Based on the 'true alignment' (center-left panel), the candidate model
with the highest marginal likelihood is determined, from a set of 
15 models. The set of models consists of all combinations
of all 4 nucleotides substitution models (JC69, HKY, TN93, GTR),
all 2 clock models (strict and relaxed log-normal) and 2 birth-death
models (Yule and Birth-Death), except the inference model used as the
generative model (JC69, strict clock, Yule). The inference model that had the
highest evidence (as shown in Table \ref{tab:true_evidence}) was 
the inference model with a JC69 nucleotide substitution model,
a relaxed log-normal clock model and a Yule tree model.
The resulting posterior trees are shown in the second panel
of the third row of posteriors in figure \ref{fig:pipeline}.

From this 'best true' posterior, the difference between each of its trees is
compared to the 'true tree' (top-left panel), using the nLTT statistic,
resulting in the second error distribution in the bottom row
of figure \ref{fig:pipeline}.

@@

From this 'twin tree', a 'twin alignment' was simulated, using
the JC69 nucleotide substitution model and a strict clock model.
The resulting alignment is shown in the center-right panel
of figure \ref{fig:pipeline}.

From the 'twin alignment' the generative inference model is run as well.
Also here, the default BEAST2 inference model is used, 
which assumes a JC69 nucleotide
substitution model, a strict clock model and a Yule tree model.
The resulting posterior trees are shown in the
third panel of the third row of figure \ref{fig:pipeline}.
From this 'generative twin' posterior, 
the difference between each of its trees is
compared to the 'twin tree' (top-right panel), using the nLTT statistic,
resulting in the error distribution shown in the third panel
of the bottom row of figure \ref{fig:pipeline}.

Based on the 'twin alignment' (center-right panel), the candidate model
with the highest marginal likelihood is determined, from the same 
set of 15 candidate models.The inference model that had the
highest evidence (as shown in Table \ref{tab:twin_evidence}) was 
the inference model with a JC69 nucleotide substitution model,
a strict clock model and a BD tree model (note: this indeed matches how
the twin tree and twin tree were simulated).
From the 'twin alignment' this best candidate inference model is run.
The resulting posterior trees are shown in the fourth panel
of the third row of posteriors in figure \ref{fig:pipeline}.

From this 'best twin' posterior (fourth in third row of
figure \ref{fig:pipeline}), the difference between each of its trees was
compared to the 'twin tree' (top-right panel), using the nLTT statistic,
resulting in the fourth error distribution in the bottom row
of figure \ref{fig:pipeline}.

@@

\new{
  From figures \ref{fig:example_500_nucleotides}, 
  \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} 
  we can observe that the discrepancy between the true and twin error 
  distributions tends to become smaller as the number of nucleotides increase (see also Fig. \ref{fig:error_vs_seqlength}).
  This occurred for both the generative and best candidate cases.
  This follows the expectation that a prior becomes less important when
  more information becomes available.
}

@@

\new{
  Because the two models are very similar to each other (the BD model can be 
  turned into a Yule model just by setting the extinction parameter to 
  zero [\cite{nee1994reconstructed}]) the median discrepancy is
  almost negligible. 
  However, with respect to the previous 
  case (subsection \ref{subsec:simplest_correct_parameterization}), 
  where a Yule tree prior was used, 
  the distributions here exhibit a greater difference.
}
\new{As we use only extant trees, it is reasonable that the method
  is slightly weaker in distinguishing between the Yule and BD
  models. It is unknown what the discriminatory power would be
  when comparing trees with extinction events.}

@@

\new{
  Here the median errors are similar in the three settings and similar to 
  ones relative to the full dataset of \ref{subsec:distribution}. We can also 
  notice that in the case of median likelihood, the twin median error appears 
  to be lower than the true mean error. This is usually a sign that the number 
  of replicates (in this case 10) is too low to allow us to draw precise 
  conclusions from this test. We did not explore further in this 
  direction using more simulations because computational times turned to be 
  extremely high. 
  The entire run took 120 hours in total.
}
